---
title: "Survival supplement v3"
author: "Dan Park and Sachin Shah"
date: "2024-08-05"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# v3 R&R
knitr::opts_chunk$set(echo = FALSE)
library(haven)
library(sas7bdat)
library(tidyverse)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(gt)

rm(list=ls())
load("~/real lasso v13-1 2024-04-30 no ICD.RData")

VITAL <- temp2

# format date types
VITAL <- VITAL %>%
  mutate(
    across(
      .cols = c(
        end_date, 
        date_of_death, 
        af_dx_date, 
        fstvisit_date
      ),
      .fns = ~ as.Date(., format = "%Y-%m-%d")
    )
  )

# include af_dx_date to conduct a survival analysis
VITAL <- VITAL %>%
  rowwise() %>%
  mutate(
    min_date = min(end_date, af_dx_date, date_of_death, na.rm = TRUE),
    follow_time_days = as.numeric(min_date - fstvisit_date),
    follow_time = as.numeric((min_date - fstvisit_date) / 365.25)
  )  

VITAL$group <- factor(VITAL$group, levels = c("I", "C"), labels = c("Screening", "Usual care"))

```

**Supplemental Methods 1**
We used a survival model with time-to-atrial fibrillation as the outcome to test the robustness of the modified Poisson regression model used in the primary analysis. The analysis demonstrated no difference in time-to-atrial fibrillation diagnosis by randomization arm supporting the primary modeling approach. The Kaplan-Meier curve demonstrates no statistical or clinically meaningful difference in time-to-AF diagnosis.

```{r cars, echo = FALSE}
surv.fit <- survfit2(
  formula = Surv(follow_time_days, outcome) ~ group, 
  data = VITAL
)

# log-rank test
test.logrank <- survdiff(
  formula = Surv(follow_time_days, outcome) ~ group, 
  data = VITAL,
  rho = 0
)

#test.logrank$chisq
pval <- test.logrank$pvalue


plot.km <- surv.fit %>%
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Percent free from AF diagnosis"
  ) +
  add_confidence_interval() + 
  # scale_color_discrete(
  #   labels = c("Screening arm", "Control arm")
  # ) + 
  # scale_fill_discrete(
  #   labels = c("Screening arm", "Control arm")
  # ) +
  scale_x_continuous(
    name = "Days following initial visit") +
  scale_y_continuous(
    name = "Proportion free from AF diagnosis", 
    limits = c(0.95,1), 
    breaks = c(seq(from = 0.95, to =1, by = .01))
    ) +
  annotate(
    "text", 
    x = 300, 
    y=0.995, 
    hjust = 0,  
    size=4, 
    label = paste(
      "p value =", 
      round(
        pval, 2)
      )) + 
  add_risktable(
    risktable_stats = "n.risk",
    stats_label = list(n.risk = "Number at Risk"),
  ) +
  theme(legend.position="top")

ggsave(
  filename = "KM_AF_dx_risk.pdf",
  plot = plot.km,
  #device = "pdf",
  path = "Tables\ and\ Figures", 
  width = 6, 
  height = 6
)

ggsave(
  filename = "KM_AF_dx_risk.png",
  plot = plot.km,
  #device = "pdf",
  path = "Tables\ and\ Figures", 
  width = 6, 
  height = 6
)
```

Further, among those with a new diagnosis of AF during the study follow up, there was no statistical or clinically meaningful difference in the distribution of time to AF diagnosis. 
```{r, echo= F}
tbl <- VITAL %>% 
  filter(outcome == 1) %>%
  select(follow_time_days, group) %>%
  tbl_summary(
    by = group,
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c(
        "{mean} ({sd})",
        "{median} ({p25}, {p75})"
    ),
    label = follow_time_days ~ "Days since first visit"
  ) %>%
  add_stat(
    fns = everything() ~ function(data, variable, by, ...) {
      p.value.t <- t.test(
        formula = as.formula(paste(variable, by, sep = "~")),
        data = data,
        var.equal = TRUE
      )$p.value
      
      p.value.wilcox <- wilcox.test(
        formula = as.formula(paste(variable, by, sep = "~")),
        data = data
      )$p.value
      
      data.frame(p.value = c(p.value.t, p.value.wilcox))
    },
    location = everything() ~ "level"
  ) %>%
  bold_p()

tbl_hnames <- show_header_names(tbl)

tbl <- tbl %>%
  modify_header(
    stat_1 = sub(
      "\\*\\*.+\\*\\*", 
      "\\*\\*Control arm\\*\\*", 
      tbl_hnames$label[which(tbl_hnames$column == "stat_1")]),
    stat_2 = sub(
      "\\*\\*.+\\*\\*", 
      "\\*\\*Screening arm\\*\\*", 
      tbl_hnames$label[which(tbl_hnames$column == "stat_2")]),
    p.value = "**p-value**"
  ) %>%
  modify_fmt_fun(
    update = p.value ~ function(x) style_pvalue(x, digits = 3),
     )
  

tbl.gt <- as_gt(tbl) %>%
  tab_footnote(
    footnote = "pooled t-test",
    locations = cells_body(
      columns = p.value,
      rows = 2
    )
  ) %>%
  tab_footnote(
    footnote = "Wilcoxon rank sum test",
    locations = cells_body(
      columns = p.value,
      row = 3
    )
  )

tbl


```

